### 2019-08-20 paper/ms.tex
initializing and eleanor LC

\documentclass[modern]{aastex62}

\include{preamble}

\shorttitle{Working Title}
\shortauthors{Bedell et al.}

\setlength{\parindent}{1.4em} % trust in Hogg
\begin{document}\sloppy\sloppypar\raggedbottom\frenchspacing % trust in Hogg

\graphicspath{ {figures/} }
\DeclareGraphicsExtensions{.pdf,.eps,.png}

\title{Title}

\author[0000-0001-9907-7742]{Megan Bedell}
\affiliation{\flatiron}

\author{Dan Foreman-Mackey}
\affiliation{\flatiron}

\author{Chelsea Huang}
\affiliation{\MIT}

\author{Jennifer Burt}
\affiliation{\MIT}

\author{Jorge Melendez}
\affiliation{\USP}

\author{Jacob L. Bean}
\affiliation{\chicago}

\author{\todo{other STPS folks}}

\correspondingauthor{Megan Bedell}
\email{mbedell@flatironinstitute.org}

\begin{abstract}\noindent
\todo{to be written}
% context
% aims
% methods
% results
\end{abstract}

\keywords{
something
---
science
}

\section{Introduction}
\label{s:intro}

\acknowledgements
We thank ...

\software{
%    \code{Astropy} \citep{astropy},
    \code{exoplanet} \citep{exoplanet},
    \code{IPython} \citep{ipython},
    \code{matplotlib} \citep{matplotlib},
    \code{numpy} \citep{numpy},
    \code{scipy} (\url{https://www.scipy.org/}),
}

% \facility{}

\bibliographystyle{aasjournal}
\bibliography{ms}

\end{document}