### 2019-09-24 paper/ms.tex
eccentricity calculations and paper outline

\documentclass[modern]{aastex62}

\include{preamble}
\newcommand{\hoststar}{\acronym{HIP}\ 96160\xspace}

\shorttitle{Working Title}
\shortauthors{Bedell et al.}

\setlength{\parindent}{1.4em} % trust in Hogg
\begin{document}\sloppy\sloppypar\raggedbottom\frenchspacing % trust in Hogg

\graphicspath{ {figures/} }
\DeclareGraphicsExtensions{.pdf,.eps,.png}

\title{A warm sub-Neptune transiting a Solar twin in TESS}

\author[0000-0001-9907-7742]{Megan Bedell}
\affiliation{\flatiron}

\author{Dan Foreman-Mackey}
\affiliation{\flatiron}

\author{Chelsea Huang}
\affiliation{\MIT}

\author{Jennifer Burt}
\affiliation{\MIT}

\author{Jhon Yana Galarza}
\affiliation{\USP}

\author{Jorge Melendez}
\affiliation{\USP}

\author{Lorenzo Spina}
\affiliation{Monash}

%\author{Jacob L. Bean}
%\affiliation{\chicago}

\author{\todo{other STPS folks}}

\correspondingauthor{Megan Bedell}
\email{mbedell@flatironinstitute.org}

\begin{abstract}\noindent
% Precise characterization of exoplanet host stars is critical to our understanding of their properties. 
% However, stellar spectroscopic characterization can be plagued with 
% context
% aims
% methods
% results
\end{abstract}

\keywords{
}

\section{Introduction}
\label{s:intro}

%The Transiting Exoplanet Survey Satellite (\TESS) has detected over one thousand candidate exoplanets to date. (blah blah)

%\hoststar has been studied extensively through a dedicated \RV planet search and spectroscopic abundance survey targeting solar twin stars. ...

\section(Data)
\label{s:data}

\subsection{Photometry}

(about \TESS photometry)

\subsection{Spectra}

(about \HARPS \RVs)

(about co-added spectrum - SNR etc)

\section{Analysis \& Results}
\label{s:analysis}

\subsection{Stellar Characterization}
\label{s:analysis:star}

(cite spectroscopic properties + abundances from previous work, plus Gaia DR2)
(isochrones fit - we use this mass \& radius going forward)

(attempts at seismology \& rotation period)

\subsection{Photometric Analysis}
\label{s:analysis:photometry}

\subsection{\RV Analysis}
\label{s:analysis:rvs}

\subsection{Joint Analysis}
\label{s:analysis:joint}

\section{Discussion}
\label{s:discussion}

(Low probability of background contaminants - discuss Gaia, spectra)

(reliability of the RV fit - potential for more planets? activity?)

(M-R diagram)

(prospects for further compositional analysis)

(prospects for atmospheric characterization)

\section{Conclusion}
\label{s:conclusion}

\acknowledgements
We thank ...

\software{
    \code{Astropy} \citep{astropy},
    \code{exoplanet} \citep{exoplanet},
    \code{IPython} \citep{ipython},
    \code{matplotlib} \citep{matplotlib},
    \code{numpy} \citep{numpy},
    \code{scipy} (\url{https://www.scipy.org/}),
}

\facility{TESS, HARPS}

\bibliographystyle{aasjournal}
\bibliography{ms}

\end{document}