### 2019-11-05 paper/ms.tex

\documentclass[modern]{aastex62}

\include{preamble}
\graphicspath{ {./figures/} }
\newcommand{\hoststar}{\acronym{HIP}\ 96160\xspace}
\newcommand{\plname}{\acronym{HIP}\ 96160b\xspace}
\newcommand{\plmass}{$13.1 \pm 5.0\ \mearth$\xspace}
\newcommand{\plradius}{$3.41 \pm 0.01\ \rearth$\xspace}
\newcommand{\pldensity}{\todo{TODO} \gcm\xspace}
\newcommand{\stmass}{$1.0 \pm x\ \msun$\xspace}
\newcommand{\stradius}{$1.0 \pm x\ \rsun$\xspace}

\shorttitle{Warm Sub-Neptune Transiting a Solar Twin}
\shortauthors{Bedell et al.}

\setlength{\parindent}{1.4em} % trust in Hogg
\begin{document}\sloppy\sloppypar\raggedbottom\frenchspacing % trust in Hogg

\graphicspath{ {figures/} }
\DeclareGraphicsExtensions{.pdf,.eps,.png}

\title{A Warm Sub-Neptune Transiting a Solar Twin in TESS}

\author[0000-0001-9907-7742]{Megan Bedell}
\affiliation{\flatiron}

\author{Dan Foreman-Mackey}
\affiliation{\flatiron}

\author{Jorge Melendez}
\affiliation{\USP}

\author{Lorenzo Spina}
\affiliation{Monash}

\author{Chelsea Huang}
\affiliation{\MIT}

\author{Jennifer Burt}
\affiliation{\MIT}

\author{Jhon Yana Galarza}
\affiliation{\USP}

%\author{Jacob L. Bean}
%\affiliation{\chicago}

\author{\todo{other STPS folks}}

\correspondingauthor{Megan Bedell}
\email{mbedell@flatironinstitute.org}

\begin{abstract}\noindent
We present the discovery of \plname, a sub-Neptune planet transiting a nearby solar twin star. 
The host star has been observed extensively as part of a radial velocity (\RV) campaign targeting solar twins with \HARPS. 
From these data, we find a planetary minimum mass of \plmass. 
We also measure a radius of \plradius from \TESS photometry. 
Taken together, the resulting planetary bulk density of \pldensity implies that an extended atmosphere is present, making this system an excellent candidate for transmission spectroscopic follow-up. 
The extreme similarity between \hoststar and the Sun means that both the absolute planetary mass and radius and the bulk composition of the \hoststar system can be measured with unusually high accuracy. 
We discuss the role of this system as a benchmark comparison to our own Solar System.

%\todo{We use the composition of the host star to place constraints on the planet's potential properties and show that it is likely to be... ??}

% Precise characterization of exoplanet host stars is critical to our understanding of their properties. 
% However, stellar spectroscopic characterization can be plagued with 
\end{abstract}

\section{Introduction}
\label{s:intro}

The Transiting Exoplanet Survey Satellite (\TESS) has detected over one thousand candidate exoplanets to date. 
The majority of these newly discovered planets orbit bright, nearby stars, making them perfectly suited for follow-up observations, including mass measurements and atmospheric spectroscopy. 
Unlike most \kepler systems, then, observational signal-to-noise may not be the dominant term in the error budget for the planet masses and other properties measured by ground-based follow-up campaigns.

(importance of solar twins)

In this work, we present the first planet detected by \TESS to orbit a true solar twin star. 
\object[HIP96160]{\hoststar} is a \stmass, \stradius G2V star with a spectrum nearly identical to that of the Sun. 
Previous studies have derived stellar properties and photospheric abundances for this \hoststar that are accurate to the $1\text{--}2\%$ level.

%All

%\hoststar has been studied extensively through a dedicated \RV planet search and spectroscopic abundance survey targeting solar twin stars. ...

\section{Data}
\label{s:data}

\subsection{Photometry}

\hoststar was observed at 2-minute cadence for a duration of 27 days in \TESS's Sector 13 pointing. 
We used the \texttt{PDCSAP\_FLUX} timeseries generated by the \TESS pipeline and available for download via the Mikulski Archive for Space Telescopes (\acronym{MAST}). 
These data have been corrected for instrumental systematic effects using an adapted version of the \kepler Presearch Data Conditioning algorithm (\acronym{PDC}; \todo{Smith12, Stumpe12,14}). 
To download the data and for some initial processing, we used the built-in routines of the \texttt{lightkurve} package \todo{(cite)}. 
This initial processing consisted of normalizing by dividing out the median flux and clipping any outliers above a limit of $+5\sigma$. 
The resulting normalized lightcurve was used as input for the model described in Section \ref{s:analysis:photometry}.
%(say some things about quality diagnostics)

\subsection{Spectra}

\hoststar was observed 56 times by the High Accuracy Radial velocity Planet Searcher (\HARPS) between 2011 and 2019. 
The bulk of these observations were made as part of a dedicated blind planet search targeting solar twins (P.I. Mel\'endez). 
All observations were carried out in high-accuracy mode with a spectral resolution R~$\sim$~115,000. 
The median \SNR is 108 pix$^{-1}$ \todo{at 600 nm}.

We adopted the \RV measurements extracted by the standard \HARPS pipeline using a cross-correlation technique with a solar-type mask \citep{Pepe2002}. 
The cross-correlation carried out by the pipeline yields additional diagnostics including the inverse bisector span (\BIS) and full width at half-maximum (\FWHM) for the line profile of the average spectral features (as selected and weighted by the mask). 
These diagnostics are commonly used as stellar activity tracers, since they quantify the line distortions which mimic Doppler shifts. 
We use these measures as furnished by the standard pipeline throughout the proceeding analysis.

In addition to these activity indicators, we also derive the \shk measurement, which quantifies the strength of emission in the cores of the Ca~\II~H\&K lines.
These were measured and corrected to the standard Mount Wilson scaling using the procedure outlined in \citet{Lovis2009}. 
Measured \shk values and photon-noise-based uncertainties are given along with the pipeline values of \RV, \BIS, and \FWHM in Table \ref{tbl:rvs}.

We dropped one observation from the analysis because its \BIS and \FWHM measurements were significant outliers from the general distribution, pointing to potential issues with the data reduction and \RV extraction.

\subsection{Gaia}

We queried Gaia Data Release 2 \citep{gaia} at the sky position and approximate magnitude of \hoststar. 
The resulting cross-match, Gaia DR2 source ID 6641996571978861440, has an estimated parallax distance of \todo{zzz} from \citet{BailerJones}, consistent with the Hipparcos estimate. 
Other \gaia information is given in Table \ref{tbl:star}. 
No additional sources were found within 30 arcseconds and 9 magnitudes of the primary match, indicating that the \TESS photometric aperture for \hoststar is unlikely to be contaminated by any major background sources.

\section{Analysis \& Results}
\label{s:analysis}

\subsection{Stellar Characterization}
\label{s:analysis:star}

(cite spectroscopic properties + abundances from previous work, plus Gaia DR2)
(isochrones fit - we use this mass \& radius going forward)

(attempts at seismology \& rotation period)

\subsection{Photometric Analysis}
\label{s:analysis:photometry}

(transits visible by eye, TOI flagging)

(centroid analysis?)

(GP model \& discussion of physical meaning - is this rotation period?)

(results of planet-only fit)

\begin{figure}
    \centering
    \includegraphics[width=\textwidth]{lightcurve.pdf}
    \caption{Photometric lightcurve of \hoststar as observed by \TESS in Sector 13. The \TESS data were fit with a Gaussian Process model (top panel) as a means of detrending. A transit model was then fit (middle panel), leaving a featureless white-noise lightcurve in the residuals (bottom panel).}
    \label{fig:lightcurve}
\end{figure}

\subsection{\RV Analysis}
\label{s:analysis:rvs}

The \RVs were first inspected for obvious periodic signals or correlations with known stellar activity indicators to inform our choice of model. 
The \HARPS\ optical fibers were replaced as part of a major instrument upgrade in June 2015, leaving an effective offset between \RVs and other line profile-sensitive components measured before and after the upgrade \citep{Lovis15}. 
For the initial inspection, we simply subtract off an offset of 0.01 fit by eye from the post-upgrade \BIS and \FWHM measurements. 
Similarly, we subtract an offset of 15.4 \ms from the post-upgrade \RVs. This quantity was derived from the median behavior of a set of \RV-quiet solar twins \citep{Melendez17}.

%\begin{figure}
%    \centering
%    \includegraphics[width=\textwidth]{periodograms.pdf}
%    \caption{Lomb-Scargle periodograms of the \HARPS\ \RV, BIS, FWHM, and \shk timeseries measurements. Horizontal lines indicate the 5\%, 1\%, and 0.1\% false alarm probability levels. Vertical red lines mark the period of the \TESS-detected planet. The \RV, BIS, and FWHM measurements used here have been corrected for the \HARPS upgrade offset.}
%    \label{fig:periodograms}
%\end{figure}

A Lomb-Scargle periodogram of the upgrade-corrected \RVs does not reveal an obvious excess of strength near the 17-day period of the \TESS planet (Figure \ref{fig:periodograms}). 
The most significant peak is instead near \todo{30} days. 
A similar peak dominates the \FWHM periodogram, suggesting that this signal may originate in stellar variability. 
Indeed, a \todo{30}-day rotation period is reasonable for this solar twin of approximately solar age.
\todo{(discuss \BIS, \shk)}

As suggested by the above periodogram analysis, there is a significant linear correlation between \RV and \FWHM in the data, with a Pearson correlation coefficient of \todo{0.7} and a bootstrapped false alarm probability of \todo{0.001}. 
Neither of the other activity indicators show similarly strong correlations. 
For this reason, we include a linear dependence on \FWHM in the following models for the \RV signal.

The rest of our analysis was carried out within the framework of \exoplanet, an open-source \python package built on \pymc and designed to perform efficient multi-parameter optimizations and Markov Chain Monte Carlo (\mcmc) samplings for Keplerian models \citep{exoplanet}. 
We first fit a model consisting only of a second-order polynomial \RV trend and a linear correlation with \FWHM. 
The free parameters for this model include coefficients of the polynomial and correlation terms, but also instrument upgrade offsets for both the \RVs

(adding a planet to the model)

(evidence for additional periods?)

\subsection{Joint Analysis}
\label{s:analysis:joint}

\begin{figure}
    \centering
    \includegraphics[width=\textwidth]{transits.pdf}
    \caption{De-trended lightcurve data and best-fit model for the two transit events in \TESS.}
    \label{fig:transits}
\end{figure}

\begin{figure}
    \centering
    \includegraphics[width=16cm]{rvphased_1pl_wfwhm}
    \caption{Phase-folded radial velocities for \plname. Individual data points and their photon-noise-based uncertainties are shown as black points and error bars, while the grey error bars represent the uncertainties inflated by the best-fit (posterior median) jitter parameters. Green points are the error-weighted means within a series of phase bins. The best-fit model is shown as a solid orange line, with the shaded region around it marking the model's 1-$\sigma$ credible interval.}
    \label{fig:rv_phased}
\end{figure}

\section{Discussion}
\label{s:discussion}

(Low probability of background contaminants - discuss Gaia, spectra)

(reliability of the RV fit - potential for more planets? activity?)

(M-R diagram)

(prospects for further compositional analysis)

(prospects for atmospheric characterization)

(star \& Sun are very similar but ended up with very different planetary systems. why? bring up [lack of] giant planet companions, cite Bryan; if migration happened, it left no trace on the star. also cite Kepler-11 as another similar star to the Sun with a different planetary system architecture.)

\section{Conclusion}
\label{s:conclusion}

\acknowledgements
We thank ... 
This work is based on observations collected at the European Southern Observatory under ESO programmes 188.C-0265 and 0100.D-0444. 
This work has made use of data from the European Space Agency (ESA) mission
{\it Gaia} (\url{https://www.cosmos.esa.int/gaia}), processed by the {\it Gaia}
Data Processing and Analysis Consortium (DPAC,
\url{https://www.cosmos.esa.int/web/gaia/dpac/consortium}). Funding for the DPAC
has been provided by national institutions, in particular the institutions
participating in the {\it Gaia} Multilateral Agreement.

\software{
    \code{Astropy} \citep{astropy},
    \code{exoplanet} \citep{exoplanet},
    \code{IPython} \citep{ipython},
    \code{matplotlib} \citep{matplotlib},
    \code{numpy} \citep{numpy},
    \code{scipy} (\url{https://www.scipy.org/}),
}

\facility{TESS, HARPS}

\bibliographystyle{aasjournal}
\bibliography{ms}

\end{document}